## PRACTICAL MACHINE LEARNING: PROJECT

### ABP

### 24 SEPTEMBER 2015

======================================================================


## Data

The data comes from research on Human Activity Recognition. Specifically, it comes from research culminating into a 2013 paper titled "Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements" by Ugulino, Cardigor, et al. In this study, six males aged 20-28 performed dumbbell curls with sensors that measured how well they performed according to guidance from trainers. Each subject completed one set of 10 repetitions with a 1.25 kg dumbbell. If the sensors indicated they lifted as taught, the performance received an "A". If they threw their elbows
to the front, they received a "B". If they lifted the dumbbell only halfway, they received a "C". Lowering the dumbbell only halfway resulted in a "D", while throwing the hips to the front resulted in an "E".


```{r}
library(data.table)
library(caret)
library(e1071)

data <- read.csv("pml-training.csv", stringsAsFactors=FALSE)


# Restrict variables to those that (1) are not missing values AND (2) related to the 
# forearm, arm, dumbbell, and belt.

missing <- sapply(data, function (x) any(is.na(x) | x == ""))
poss <- !missing & grepl("forearm|[^fore)]arm|dumbbell|belt", names(missing))

maybe <- names(missing)[poss]

# Restrict variables to only (1) the possible candidates AND (2) the dependent 
# variable, "classe".

allvar <- c("classe", maybe)

data2 <- data[,allvar]

# Convert the data set into a data table and "classe" into a factor variable.

data_dt <- data.table(data2)

data3 <- data_dt[, classe := factor(data_dt[,classe])]

data3[, .N, classe]

# Partition the data.

set.seed(5150)
train <- createDataPartition(data3$classe, p=0.7)
train_a <- data3[train[[1]]]
train_b <- data3[-train[[1]]]


# Center and scale both data sets.

cs <- train_a[,maybe, with=FALSE]
pp <- preProcess(cs)

pp_cs <- predict(pp, cs)
train_a_cs <- data.table(data.frame(classe = train_a[,classe],pp_cs))

cs <- train_b[,maybe, with=FALSE]
pp <- preProcess(cs)

pp_cs <- predict(pp, cs)
train_b_cs <- data.table(data.frame(classe = train_b[,classe],pp_cs))

# Set control parameters.

ctrlPar <- trainControl(classProbs=TRUE, savePredictions=TRUE)

```


## Methodology

This analysis will use random forest to make predictions. We
will use Random Forest on the training set A. If all goes
well, the out of sample error rate should be approximately
2%.


```{r cache=TRUE}

# Use Random Forest ensemble on training set A.

trainingMod <- train(classe ~ ., data=train_a_cs,
method="rf")

# Evaluate the data with the training set A, which has 70%
of the original training data.

hat <- predict(trainingMod, train_a_cs)
confusionMatrix(hat, train_a[,classe])

# Evaluate the data with the training set B, which has 30%
of the original training data.

hat <- predict(trainingMod, train_b_cs)
confusionMatrix(hat, train_b[,classe])


varImp(trainingMod)
trainingMod$finalModel

```


The results show that the error rate is less than 1%.

Now we will use the testing data to make predictions.  With
that, we combine the predictions with the appropriate fields
and display the results. When feeding the results into
Coursera's Prediction Assignment Submission, the results
were 20 for 20, a 100% success rate.

```{r}

# Compare with the testing data. First, prep the testing
data.

t_data <- read.csv("pml-testing.csv",
stringsAsFactors=FALSE)

t_data_cs <- predict(pp, t_data[, maybe])
hat <- predict(trainingMod, t_data_cs)

t_data <- cbind(hat, t_data)
t_data2 <- subset(t_data,
select=names(t_data)[grep("forearm|[^fore)]arm|dumbbell|belt
", names(t_data), invert=TRUE)])

# Display final predictions.

t_data2

```


## Sources

Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu,
R.; Fuks, H. Wearable Computing: Accelerometers' Data
Classification of Body Postures and Movements. Proceedings
of 21st Brazilian Symposium on Artificial Intelligence.
Advances in Artificial Intelligence - SBIA 2012. In: Lecture
Notes in Computer Science. , pp. 52-61. Curitiba, PR:
Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9.
DOI: 10.1007/978-3-642-34459-6_6.
